Almost compatible functions and infinite length games

### 2015-12-03 articles/almost-compatible-functions-and-topological-games/article.tex
almost-complete infinite length game applications

  \begin{question}
    Does there exist a topological space \(X\) for which
    \(\pl F\win\menGame{X}\) but \(\pl F\notkmarkwin{2}\menGame{X}\)?
    (That is, the second player can win the Menger game on \(X\)
    with perfect information but not with \(2\)-Markov information.)
  \end{question}

  One might hope that \(\oneptlind{(\mf c^+)}\) might answer this question
  in the affirmative as \(\alcompS{\mf c^+}\) fails, but we will show that assuming
  \(V=L\), \(\pl F\kmarkwin{2}\menGame{\oneptlind\kappa}\) for all
  cardinals \(\kappa\).

@@

  \section{More consistency results}

@@

  We are also able to force \(\alcompSP{\kappa}\) to fail for some cardinal.

@@

  \newpage

  \section{TODO ALL THIS STUFF NEEDS EDITING STILL}

  \section*{Applications!}

@@

    S'(\kappa) &
    \pl F \kmarkwin2\schFillInitialGame\kappa &
    \pl F \kmarkwin2\schFillGame\kappa \\

    &
    &
    \pl F \ktactwin2\schFillStrictGame\kappa \\
  };
  \path[>=latex,->]
    (m-2-1) edge (m-1-2)
    (m-2-1) edge (m-2-2)
    (m-2-1) edge (m-3-3)
    (m-1-2) edge (m-2-2)
    (m-1-3) edge (m-2-3)
    (m-3-3) edge (m-2-3);
  \path[>=latex,<->]
    (m-1-2) edge (m-1-3)
    (m-2-2) edge (m-2-3);
\end{tikzpicture}
\end{center}
\caption{Diagram of Scheeper/Menger game implications with \(\alcompSP{\kappa}\)}
\label{GamesDiagram2}
\end{figure}

  \begin{theorem}[\cite{clontzMengerGamePreprint}]
    Figure \ref{GamesDiagram2} holds.
    (Actually, TODO double-check that it works with just S', particularly
    the strict game)
  \end{theorem}

  It was left open if these implications can be reversed. The answer is
  consistently no.

@@

  \begin{proof}
    Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    \(\schFillInitialGame{\omega_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    for \(\pl F\) in \(\schFillInitialGame{\omega_\alpha}\) as follows:
    \[
      \sigma(\<C\>,0)
        =
      \sigma_0(\<C\cap\omega_{\beta_0}\>,0)
    \]
    \[
      \sigma(\<C,D\>,n+1)
        =
      \sigma_{n+1}(\<D\cap\omega_{\beta_{n+1}}\>,0)
        \cup
      \bigcup_{m\leq n}
      \sigma_m(\<C\cap\omega_{\beta_m},D\cap\omega_{\beta_m}\>,n-m+1)
    \]

    Let \(\<C_0,C_1,\dots\>\) be an attack by
    \(\pl C\) in \(\schFillInitialGame{\omega_\alpha}\), and
    \(\alpha\in C_0\).
    Choose \(N<\omega\) with \(\alpha<\omega_{\beta_{N+1}}\). Consider the
    attack
    \(\<C_{N+1}\cap\omega_{\beta_{N+1}},C_{N+2}\cap\omega_{\beta_{N+1}},\dots\>\)
    by \(\pl C\) in \(\schFillInitialGame{\omega_{\beta_{N+1}}}\). Since
    \(\sigma_{N+1}\) is a winning strategy and
    \(\alpha\in C_{N+1}\cap\omega_{\beta_{N+1}}\), either
    \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\omega_{\beta_{N+1}}\>,0)\) and thus
    \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    \(
      \alpha
        \in
      \sigma_{N+1}(
        \<C_{N+M+1}\cap\omega_{\beta_{N+1}},C_{N+M+2}\cap\omega_{\beta_{N+1}}\>
        ,M+1
      )
    \)
    for some \(M<\omega\) and thus
    \(
      \alpha
        \in
      \sigma(
        \<C_{N+M+1},C_{N+M+2}\>,
        N+M+2
      )
    \). Thus \(\sigma\) is a winning strategy.
  \end{proof}

  \begin{corollary}
    It is consistent that \(\alcompSP{\omega_\omega}\) fails,
    but as \(\alcompSP{\omega_k}\) holds for all \(k<\omega\), we have
    \(\pl F\kmarkwin2\schFillIntGame{\omega_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\omega_\omega}\).
  \end{corollary}

  A tricky topological question: does \(\pl F\win\menGame{X}\) imply
  \(\pl F\kmarkwin2\menGame{X}\)?
  (C showed that )
  Under \(V=L\), we cannot hope to find
  a counterexample using \(X=\oneptlind\kappa\) since
  \(\alcompSP{\kappa}\) and thus \(\pl F\kmarkwin2\schFillIntGame{\kappa}\)
  always holds.

  \begin{definition}
    Let \(R_\omega\) be the real numbers with the topology of the usual
    open intervals with countably many elements removed.
  \end{definition}

  \begin{theorem}
    \(\pl F\win\menGame{R_\omega}\).
    If there exists a Kurepa family on the reals, then
    \(\pl F\kmarkwin2\menGame{R_\omega}\).
  \end{theorem}

>>>>

  \begin{question}\label{mainQuestion}
    Does there exist a topological space \(X\) for which
    \(\pl F\win\menGame{X}\) but \(\pl F\notkmarkwin{2}\menGame{X}\)?
    (That is, the second player can win the Menger game on \(X\)
    with perfect information but not with \(2\)-Markov information.)
  \end{question}

@@

  \section{Consistency results}

@@

  We are also able to force \(\alcompSP{\kappa}\) to fail for every cardinal
  other than the first \(\omega\)-many we've already guaranteed.

@@

  \section{Applications to infinite length games}

  We introduce three variations of Scheeper's game which we defined in
  the introduction.

  \begin{game}
    Let \(\schFillGame\kappa\) denote the
    \term{Scheepers countable-finite union game} which proceeds analogously
    to \(\schFillStrictGame\kappa\), except that \(\pl C\)'s restriction in round \(n+1\)
    is weakened to \(C_{n+1}\supseteq C_n\).
  \end{game}

  \begin{game}
    Let \(\schFillInitialGame\kappa\) denote the
    \term{Scheepers countable-finite initial game} which proceeds analogously
    to \(\schFillGame\kappa\), except that \(\pl F\)'s winning condition is
    weakened to \(\bigcup_{n<\omega}F_n\supseteq C_0\).
  \end{game}

  \begin{game}
    Let \(\schFillIntGame\kappa\) denote the
    \term{Scheepers countable-finite intersection game} which proceeds analogously
    to \(\schFillInitialGame\kappa\), except that \(\pl C\) may choose any
    \(C_n\in[\kappa]^{\leq\omega}\) each round, and \(\pl F\)'s winning condition
    is weakened to
    \(\bigcup_{n<\omega}F_n\supseteq\bigcap_{n<\omega}C_n\).
  \end{game}

  In \cite{clontzMengerGamePreprint} Clontz extended Scheepers' application of
  almost-compatible injections to these game variants as well as
  \(\menGame{\oneptlind\kappa}\).

@@

    \alcompSP\kappa \\

    &
    \pl F \kmarkwin2\schFillInitialGame\kappa &
    \pl F \kmarkwin2\schFillGame\kappa \\
  };
  \path[>=latex,->]
    (m-2-1) edge (m-1-2)
    (m-2-1) edge (m-3-2)
    (m-1-2) edge (m-3-2)
    (m-1-3) edge (m-3-3);
  \path[>=latex,<->]
    (m-1-2) edge (m-1-3)
    (m-3-2) edge (m-3-3);
\end{tikzpicture}
\end{center}
\caption{Diagram of Scheeper/Menger game implications with \(\alcompSP{\kappa}\)}
\label{GamesDiagram2}
\end{figure}

%   \begin{figure}[h]
% \begin{center}
% \begin{tikzpicture}
%   \matrix (m) [matrix of math nodes,row sep=3em,column sep=1em,minimum width=2em]
%   {
%     &
%     \pl F \kmarkwin2\menGame{\oneptlind\kappa} &
%     \pl F \kmarkwin2\schFillIntGame\kappa \\

%     \alcompSP\kappa &
%     \pl F \kmarkwin2\schFillInitialGame\kappa &
%     \pl F \kmarkwin2\schFillGame\kappa \\

%     &
%     &
%     \pl F \ktactwin2\schFillStrictGame\kappa \\
%   };
%   \path[>=latex,->]
%     (m-2-1) edge (m-1-2)
%     (m-2-1) edge (m-2-2)
%     (m-2-1) edge (m-3-3)
%     (m-1-2) edge (m-2-2)
%     (m-1-3) edge (m-2-3)
%     (m-3-3) edge (m-2-3);
%   \path[>=latex,<->]
%     (m-1-2) edge (m-1-3)
%     (m-2-2) edge (m-2-3);
% \end{tikzpicture}
% \end{center}
% \caption{Diagram of Scheeper/Menger game implications with \(\alcompSP{\kappa}\)}
% \label{GamesDiagram2}
% \end{figure}

  \begin{theorem}
    \(\alcompSP\kappa\) implies the game-theoretic conditions
    shown in Figure \ref{GamesDiagram2}.
  \end{theorem}

  \begin{proof}
    The weaker claim
    \(\alcompS\kappa\Rightarrow\pl F \kmarkwin2\schFillIntGame\kappa\)
    was proven in \cite{clontzMengerGamePreprint};
    however, the strategy
    only required that the \(f_A\) be pairwise almost-compatible and
    that preimages of finite sets in each \(f_A\)
    are finite, which is also possible assuming \(\alcompSP\kappa\).
  \end{proof}

    %TODO maybe this can be made to work...
    % In \cite{MR1129143} Scheepers used a proof of Todorcevic to demonstrate
    % \(\alcompS\kappa\Rightarrow\pl F \ktactwin2\schFillStrictGame\kappa\).
    % For the convenience of the reader we explicitly demonstrate
    % the generalization as follows.

    % Let \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) witness
    % \(\alcompSP\kappa\), and define \(g_A:A\to\omega\) by
    % \(g_A(\alpha)=f_A(\alpha)-|\{\beta\in A:f_A(\beta)<f_A(\alpha)\}|\).

    % We claim that \(\{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}\)
    % must be finite as it is bounded above by
    % \(\max\{M,f_A(\alpha),f_B(\alpha):f_A(\alpha)\not=f_B(\alpha)\}\)
    % where \(M=f_B(\alpha)\) for some \(\alpha\in B\setminus A\).
    % To see this, let \(f_A(\alpha)=f_B(\alpha)=N>M\) and assume
    % \(f_A(\beta)\not=f_B(\beta)\) implies \(f_A(\beta),f_B(\beta)<N\).
    % Then
    % \[
    %   g_A(\alpha)
    %     =
    %   N-|\{\beta\in A:f_A(\beta)<N\}|
    %     >
    %   N-|\{\beta\in B:f_B(\beta)<N\}|
    %     =
    %   g_B(\alpha)
    % \]
    % with the strictness of the inequality witnessed by \(f_B(\alpha)=M<N\)
    % for some \(\alpha\in B\setminus A\).

    % As a result,
    % \[
    %   \sigma(\<A,B\>)
    %     =
    %   \{\alpha\in A:f_A(\alpha)\not=f_B(\alpha)\}
    %     \cup
    %   \{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}
    % \]
    % is a legal \(2\)-tactic for \(\pl F\). Let \(C=\<C_0,C_1,\dots\>\) be
    % a strictly increasing sequence of countable sets and
    % \(\alpha\in C_n\). Of course if
    % \(f_{C_{n+m}}(\alpha)\not=f_{C_{n+m+1}}(\alpha)\) for some
    % \(m<\omega\) we have \(\alpha\in\sigma(\<C_{n+m},C_{n+m+1}\>)\) immediately.
    % Otherwise let \(f_{C_{n+m}}(\alpha)=f_{C_{n+m+1}}(\alpha)=N\)
    % for all \(m<\omega\).

  It was left open if \(\alcompSP\kappa\) was equivalent to these
  game-theoretic conditions. The answer is consistently negative.

@@

  \begin{proof}
    The proof proceeds nearly identically to the previous proof.
  \end{proof}
    % Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    % \(\schFillInitialGame{\omega_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    % for \(\pl F\) in \(\schFillInitialGame{\omega_\alpha}\) as follows:
    % \[
    %   \sigma(\<C\>,0)
    %     =
    %   \sigma_0(\<C\cap\omega_{\beta_0}\>,0)
    % \]
    % \[
    %   \sigma(\<C,D\>,n+1)
    %     =
    %   \sigma_{n+1}(\<D\cap\omega_{\beta_{n+1}}\>,0)
    %     \cup
    %   \bigcup_{m\leq n}
    %   \sigma_m(\<C\cap\omega_{\beta_m},D\cap\omega_{\beta_m}\>,n-m+1)
    % \]

    % Let \(\<C_0,C_1,\dots\>\) be an attack by
    % \(\pl C\) in \(\schFillInitialGame{\omega_\alpha}\), and
    % \(\alpha\in C_0\).
    % Choose \(N<\omega\) with \(\alpha<\omega_{\beta_{N+1}}\). Consider the
    % attack
    % \(\<C_{N+1}\cap\omega_{\beta_{N+1}},C_{N+2}\cap\omega_{\beta_{N+1}},\dots\>\)
    % by \(\pl C\) in \(\schFillInitialGame{\omega_{\beta_{N+1}}}\). Since
    % \(\sigma_{N+1}\) is a winning strategy and
    % \(\alpha\in C_{N+1}\cap\omega_{\beta_{N+1}}\), either
    % \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\omega_{\beta_{N+1}}\>,0)\) and thus
    % \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    % \(
    %   \alpha
    %     \in
    %   \sigma_{N+1}(
    %     \<C_{N+M+1}\cap\omega_{\beta_{N+1}},C_{N+M+2}\cap\omega_{\beta_{N+1}}\>
    %     ,M+1
    %   )
    % \)
    % for some \(M<\omega\) and thus
    % \(
    %   \alpha
    %     \in
    %   \sigma(
    %     \<C_{N+M+1},C_{N+M+2}\>,
    %     N+M+2
    %   )
    % \). Thus \(\sigma\) is a winning strategy.

  \begin{corollary}
    It is consistent that \(\alcompSP{\aleph_\omega}\) fails,
    but as \(\alcompSP{\aleph_k}\) holds for all \(k<\omega\), we have
    \(\pl F\kmarkwin2\schFillIntGame{\omega_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\omega_\omega}\).
  \end{corollary}

  Our topological question \ref{mainQuestion} remains unsolved; however,
  our results have revealed that we cannot hope to find a \(ZFC\)
  counterexample where \(X=\oneptlind\kappa\). This is
  because if we also assume \(V=L\), it follows that
  \(\alcompSP{\kappa}\) and therefore \(\pl F\kmarkwin2\menGame{X}\)
  for every cardinal.

  \section{Generalized Kurepa families and almost-compactible functions}

  TODO
